---
title: "Block feature selection with UBayFS"
author: "Anna Jenul, Stefan Schrunner"
date: "`r Sys.Date()`"
output: 
  prettydoc::html_pretty:
    theme: architect
    toc: TRUE
    hightlight: github
    geometry: width=400cm
bibliography: references.bib
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r set-options, echo=FALSE, cache=FALSE}
options(width = 10000)
```


```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
knitr::opts_chunk$set(fig.height = 5)
knitr::opts_chunk$set(fig.width = 8)
knitr::opts_chunk$set(fig.align = 'center')
```

## Intro Block feature selection
Block feature selection is essential in different application fields, such as healthcare. For diseases like cancer, multiple sources of data are usually collected, including clinical data, genetic data, or image data, where measurements from each source present a separate feature block. Imagine that not all data sources are relevant for machine learning models and data analysis. Suppose we knew that image data do not contain additional information to other feature blocks. In that case, a hospital might save a lot of resources and money by not collecting those data in advance. Block feature selection can be used for such purposes. A prominent method for block feature selection is the group Lasso [@yuan:groupLasso] which is a crossover of the lasso method on block variables. Though, UBayFS is more flexible and needs no hyperparameter estimation as in group Lasso.

## UBayFS example
At first we load the package and the Breast Cancer Wisconsin (BCW) dataset, which is described in the main vignette.
```{r, include = TRUE, cache = TRUE}
library(UBayFS)
data(bcw)
```


For block feature selection, it is necessary to define each feature's block affiliation, which can be done either (a) by a block list, or (b) directly by a matrix as input. 

### Version (a): block list
The first example demonstrates version (a). We define three blocks for the BCW dataset, where the first block contains features 1 to 10, block two features 11 to 20, and block three features 21 to 30.
```{r, include = TRUE}
block_list = list()
block_list[[1]] = c(1:10)
block_list[[2]] = c(11:20)
block_list[[3]] = c(21:30)
```

Furthermore, we define a help function to build block weights, simplifying the weight affiliation. Features from the same block get the same prior weight. In this example, features from block 1 get a weight of 0.5, features from block 2 get a weight of 1, and features from block 3 gain weight of 2.
```{r, include = TRUE}
build_block_weights = function(blocks, weights){
	return(weights[blocks])
}
build_block_weights(blocks = rep(c(1,2,3), each = 10), weights = c(0.5,1,2))
```

For the UBayFS model, we define a max size constraint with three features and a max size block constraint with maximal one block.

```{r, include = TRUE}
set.seed(1) # for reproducibility
model = build.UBaymodel(data = bcw$data,
                        target = bcw$labels,
                        M = 100, 
                        tt_split = 0.75,
                        nr_features = 10,
                        method = "mRMR",
                        prior_model ="dirichlet",
                        weights = build_block_weights(blocks = rep(c(1,2,3), each = 10),
                                                      weights = c(1,1,1)),
                        lambda = 1,
                        constraints = buildConstraints(constraint_types = c("max_size"), 
                                                       constraint_vars = list(3), 
                                                       num_elements = dim(bcw$data)[2], 
                                                       rho = 1),
                        block_constraints = buildConstraints(constraint_types = c("max_size"), 
                                                             constraint_vars = list(1), 
                                                             num_elements = 3, 
                                                             rho = 1, 
                                                             block_list = block_list),
                        optim_method = "GA",
                        popsize = 100,
                        maxiter = 100,
                        shiny = FALSE
                        )
summary(model)
```

All further steps are similar to the UBayFS variant without block constraints. In this example, three features from the third block are selected. Hence, all constraints are fulfilled.
```{r, include = TRUE}
model = train(model)
summary(model)
plot(model)
```

```{r, include = TRUE}
evaluateMultiple(state = model$output$feature_set[1,], model = model)
block_admissibility(state = unlist(model$output$feature_set[1,]), 
                    constraints = model$constraint.params$block_constraints)
```

### Version (b): block matrix
In version (b), the block structure can be a direct input which we define as an R matrix. Again, the first block contains features 1 to 10, block two features 11 to 20, and block three features 21 to 30.

```{r, include = TRUE}
block_matrix = matrix(0, nrow = 3, ncol = ncol(model$data))
block_matrix[1, 1:10] = 1
block_matrix[2, 11:20] = 1
block_matrix[3, 21:30] = 1

block_matrix
```

The rest of the evaluation is the same for version (a) and (b). Setting the same seed, we achieve the same result of selected features.
```{r, include = TRUE}
set.seed(1)
model = build.UBaymodel(data = bcw$data,
                        target = bcw$labels,
                        M = 100, 
                        tt_split = 0.75,
                        nr_features = 10,
                        method = "mRMR",
                        prior_model ="dirichlet",
                        weights = build_block_weights(blocks = rep(c(1,2,3), each = 10),
                                                      weights = c(1,1,1)),
                        lambda = 1,
                        constraints = buildConstraints(constraint_types = c("max_size"), 
                                                       constraint_vars = list(3), 
                                                       num_elements = dim(bcw$data)[2], 
                                                       rho = 1),
                        block_constraints = buildConstraints(constraint_types = c("max_size"), 
                                                             constraint_vars = list(1), 
                                                             num_elements = 3, 
                                                             rho = 1, 
                                                             block_matrix = block_matrix),
                        optim_method = "GA",
                        popsize = 100,
                        maxiter = 100,
                        shiny = FALSE
                        )
summary(model)
```


```{r, include = TRUE}
model = train(model)
summary(model)
plot(model)
```

```{r, include = TRUE}
evaluateMultiple(state = model$output$feature_set[1,], model = model)
block_admissibility(state = unlist(model$output$feature_set[1,]), 
                    constraints = model$constraint.params$block_constraints)
```


## References
